Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-10 paper.tex
Lots of improvements, especially in describing bigraphs

\newcommand\Dist{\mathit{Dist}}
\newcommand\ctrl{\mathit{ctrl}}
\newcommand\prnt{\mathit{prnt}}
\newcommand\link{\mathit{link}}
\newcommand\points{\mathit{points}}
\newcommand\ports{\mathit{ports}}
\newcommand\ar{\mathit{ar}}
\newcommand\id{\mathsf{id}}
\newcommand\Id{\mathsf{Id}}

@@

\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Bigraphs \and Bigraphical Reactive Systems \and Probabilistic
  semantics \and Markov decision process.}
\end{abstract}

@@

@@

\begin{definition}[rewards]
  A \emph{reward structure} for an MDP $(S, \overline{s}, A, P, L)$ is a pair
  $(\rho, \iota)$, where $\rho : S \to \mathbb{R}$ is the \emph{state reward
    function}, and $\iota : S \times A \to \mathbb{R}$ is the \emph{transition
    reward function}.
\end{definition}

@@

% TODO: introduce bigrapher
From \cite{dblp:journals/tcs/sevegnanic15}.
From PhD thesis \cite{DBLP:phd/ethos/Sevegnani12}

@@

Consider the graphical representation of a bigraph in Fig.
\ref{example_bigraph}. The white ellipses are the \emph{nodes}. Each node has a
type, called \emph{control}, denoted here by the labels \textsf{A}, \textsf{B},
and \textsf{C}. The two white dashed rectangles represent \emph{regions},
i.e. parents of any otherwise parentless nodes. Grey dashed rectangles are
called \emph{sites} and encode parts of the model that have been abstracted
away. Nodes, sites, and regions are also known as \emph{places}. Two places with
the same parent, or two points with the same link are called \emph{siblings}.

@@

\begin{definition}[composition for place graphs with sharing]
  If $F : k \to m$ and $G : m \to n$ are two concrete place graphs with sharing
  with $V_F \cap V_G = \emptyset$, their composite
  \[ G \circ F = (V, \ctrl, \prnt) : k \to n \]
  has nodes $V = V_F \uplus V_G$ and control map $\ctrl = \ctrl_F \uplus
  \ctrl_G$. Its parent relation $\prnt \subseteq (k \uplus V) \times (V \uplus
  n)$ is given by:
  \[ \prnt \coloneqq \prnt_G^\lhd \uplus \prnt_\circ \uplus \prnt_F^\rhd, \]
  where
  \begin{align*}
    \prnt_F^\rhd &= \prnt_F \rhd V_F, \\
    \prnt_G^\lhd &= V_G \lhd \prnt_G, \\
    \prnt_\circ &= (m \lhd \prnt_G) \circ (\prnt_F \rhd m).
  \end{align*}
\end{definition}

\begin{definition}[tensor product for place graphs]
  If $G_0 : m_0 \to n_0$ and $G_1 : m_1 \to n_1$ are two concrete place graphs
  with sharing with $V_F \cap V_G = \emptyset$, their tensor product
  \[ G_0 \otimes G_1 = (V, \ctrl, \prnt) : m_0 + m_1 \to n_0 + n_1 \]
  has nodes $V = V_{G_0} \uplus V_{G_1}$ and control map $\ctrl \coloneqq
  \ctrl_{G_0} \uplus \ctrl_{G_1}$. Its parent relation $\prnt \subseteq [(m_0 +
  m_1) \uplus V] \times [V \uplus (n_0 + n_1)]$ is defined as
  \[ \prnt_{G_0} \uplus \prnt_{G_1}^{(m_0, n_0)}, \]
  where
  \begin{align*}
    \prnt_{G_1}^{(m_0, n_0)} = &\{ (v, w) \mid (v, w) \in \prnt_{G_1} \text{ and } v, w \in V_{G_1} \} \\
    \uplus &\{ (m_0 + i, w) \mid (i, w) \in \prnt_{G_1}, w \in V_{G_1} \text{ and } i \in m_1 \} \\
    \uplus &\{ (v, n_0 + i) \mid (v, i) \in \prnt_{G_1}, v \in V_{G_1} \text{ and } i \in n_1 \} \\
    \uplus &\{ (m_0 + i, n_0 + j) \mid (i, j) \in \prnt_{G_1}, i \in m_1, \text{ and } j \in n_1 \}.
  \end{align*}
\end{definition}

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectivelly the inner and outer names of
  the link graph. $F$ has finite sets $V_F \subset \mathcal{V}$ of nodes and
  $E_F \subset \mathcal{E}$ of edges, a control map $\ctrl_F : V_F \to
  \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

The sets of points and the set of ports of a link $l$ are defined by
\[ \points_F(l) \coloneqq \{ p \mid \link_F(p) = l \}, \quad \ports_F(l)
  \coloneqq \points_F(l) \setminus X. \]
An edge is \emph{idle} if it has no points. Identities over name sets are
defined by $\id_X = (\emptyset, \emptyset, \emptyset, \Id_X) : X \to X$.

@@

Given a set of reaction rules, we refer to the configurations that a system may
adopt as \emph{states}. A bigraph with inner face $\epsilon$ is called
\emph{ground}.

@@

Lastly, BigraphER supports \emph{rule priorities}
\cite{DBLP:journals/tcs/BaetenBKW89}, i.e., a partial ordering on the reaction
rules, where a rule of lower priority can be applied only if no rule of higher
priority is applicable. Priorities are implemented by assigning each reaction
rule to a \emph{priority class}, where all classes are strictly ordered, and
reaction rules in the same class have the same priority.

\section{Extensions to Bigraphs}

@@

  Let $g, g'$ be ground bigraphs, and $\{ (r_i, r'_i, p_i) \}_{i=1}^n$ a set of
  ground probabilistic reaction rules, where for each $r_i$, there exists a
  bigraph $D_i$ such that $g = D_ir_i$. Let $S = \{ (r_i, r'_i, p_i) \mid g' =
  D_ir'_i \}$ (for the same $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation is defined as
  \[ g \xrightarrowrhd{p}_{\mathcal{R}} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', p') \in S} p'. \]
\end{definition}

@@

  Let $g, g'$ be ground bigraphs, $a \in A$ an action, and $\{ (r_i, r'_i, a,
  p_i) \}_{i=1}^n$ a set of ground nondeterministic reaction rules with action
  $a$, where for each $r_i$, there exists a bigraph $D_i$ such that $g =
  D_ir_i$. Let $S = \{ (r_i, r'_i, a, p_i) \mid g' = D_ir'_i \}$ (for the same
  $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation for action $a$ is defined as
  \[ g \xrightarrowrhd[r(a)]{p}_{a} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', a, p') \in S} p'. \]
\end{definition}

BigraphER \cite{DBLP:conf/cav/SevegnaniC16} supports defining predicates, i.e.,
bigraphs that are compared to every encountered state. We can associate a reward
with each predicate, allowing us to assign rewards to states in a flexible and
semantically meaningful way: the reward of a state is simply the sum of the
rewards of all matching predicates (and $0$ in case there are none).

@@

@@

The MDP can be represented as an NBRS with BigraphER code in Listing
\ref{example_listing}. Furthermore, the BigraphER's visualisation of the
(automatically generated) transition system is in Fig. \ref{example_ts}.

@@

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. Text inside state ellipses lists all predicates that are satisfied by
that state. Edges from actions to states are labelled with both the probability
and the name of the reaction rule. Transitions from either a fully generated
transition system or a simulation, labels from predicates, and state/transition
rewards can be exported to the corresponding PRISM plain text formats.

% TODO: introduce this
\begin{figure}
  \centering
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

@@

% TODO: Akin to \cite{dblp:conf/nfm/giaquintahimn18} and maybe \cite{koenig1998xavier}.

@@

% TODO: what is a link, mention that it's a hyperedge (at least one node)
% TODO: explain the syntax of the language
Our first scenario is concerned with collecting objects in a grid-like
two-dimensional world. The agent starts in the southwest corner of the map
(which we call \emph{home}), explores the environment until it collects a
specified number of objects, and returns home (while possibly collecting more
objects along the way). The model tracks which cells of the grid have been
explored, and only the unexplored cells have a constant probability $p$ of
rewarding the agent with an object. The desired number of objects can be easily
changed, and the map can be customised to make some transitions unavailable
(i.e. add walls), as long as the agent can still return home.

@@

We begin with a list of controls in Listing \ref{agent1_controls}. On the right
hand side of each equal sign is the number of ports that each node of that
control has, while the \texttt{\textbf{atomic}} keyword specifies the nodes that
cannot contain other nodes. \texttt{Cell}s represents discrete locations. They
are linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all
except one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}. In order to represent multiple
objects, it is convenient to put the second \texttt{Object} inside the first,
and so on. \texttt{Directions} contains a subset of the four nodes,
corresponding to the four possible directions of travel: \texttt{North},
\texttt{East}, \texttt{West}, \texttt{South}. We connect neighbouring
\texttt{Cell}s by linking the \texttt{West} node of one \texttt{Cell} to the
\texttt{East} node of another (or \texttt{North} to \texttt{South}).

@@

Next, we define two predicates: \texttt{home} and \texttt{goal} (see Listing
\ref{agent1_code}). The former states that the agent is home
if it is in the southwest corner of the grid,
while the latter specifies a goal of collecting one object. A larger number of
objects can be set by nesting \texttt{Object}s inside each other, e.g.,
\texttt{Agent.Object.Object}. Additionally, Listing \ref{agent1_code} defines
the initial $2 \times 2$ grid, as described previously.

@@

Then, we have an action for each of the four directions, each with three
reaction rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the first time and
finding an object. We will use the rules for going north in Listing
\ref{agent1_north} as an example. When going to an already visited cell, the
\texttt{Agent} simply traverses a link between the \texttt{North} node of its
previous \texttt{Cell} and the corresponding \texttt{South} node. As this is the
only reaction rule for this situation, its probability is $1$. When going to an
unvisited node, we collect an \texttt{Object} with probability $p$, or make the
transition without getting an \texttt{Object} with probability $1-p$.
Regardless, the \texttt{Cell} changes its link from \texttt{Unvisited} to
\texttt{Visited}.

@@

Lastly, we have three actions for when the agent acquires the specified number
of objects: two for going home west or south (since these are the two directions
necessary and sufficient to reach home), and one for staying home. The two
actions for going home have two reaction rules each, depending on whether the
destination cell is visited or not. They are just like the rules in Listing
\ref{agent1_north}, except every mention of \texttt{Agent} is replaced with
\texttt{goal}. The rule to stay home simply says that if the agent has achieved
the goal, and is home (characterised by a \texttt{Cell} with only two
directions: \texttt{North} and \texttt{East}), then nothing should change (see
Listing \ref{agent1_home}). To make the agent start heading home as soon as
enough objects are collected, all five reaction rules related to going and
staying home are in a higher priority class.

@@

% TODO: Must mention: \cite{DBLP:conf/giscience/WaltonW12}.

@@

In order to impement nesting, a \texttt{Cell} contains a \texttt{Node}, which
can have any number of \texttt{Cell}s of its own. Since after entering a room,
you end up in a specific place inside the room (next to the door), we also need
to link up the \texttt{Node} with one of the \texttt{Cell}s it contains (see
Fig. \ref{agent2_nesting}). Both \texttt{Cell}s and \texttt{Node}s support
having a single link for exactly this purpose. Thus, most of the \texttt{Cell}
ports are not linked to anything, which is implemented via closures (e.g.
\texttt{/x Cell\{x\}}). The initial map in Listing \ref{agent2_initial} has a
corridor of three cells, with the agent on the west-most cell, and a door to a
room in the middle cell. One moves in and out of a room via a pair of actions
with reaction rules pictured in Fig. \ref{go_in_out}.

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1}). In order for this to happen before any other action
can take place, the two reaction rules in Listing \ref{agent2_room} are put on a
higher priority class. With probability $p$, the generated room has each
\texttt{Cell} linked to its two neighbours; and with probability $1-p$, the top
two \texttt{Cell}s are not linked , possibly forcing the agent to choose a
different path. Finally, we have four actions, each with a single reaction rule
with probability $1$, corresponding to the four possible directions of travel.
The reaction rules are very similar to those of our previous example, see
Listing \ref{agent2_north}.

>>>>

\newcommand\Dist{\mathit{Dist}}
\newcommand\ctrl{\mathit{ctrl}}
\newcommand\prnt{\mathit{prnt}}
\newcommand\link{\mathit{link}}
\newcommand\ar{\mathit{ar}}
\newcommand\id{\mathsf{id}}
\newcommand\Id{\mathsf{Id}}

@@

\begin{abstract}
  In this paper we introduce two extensions of bigraphical reactive systems
  (BRSs) to handle probabilistic and nondeterministic behaviour, allowing BRSs
  to be used in discrete-time Markov chain and Markov decision process (MDP)
  generation. We extend the implementation of an open-source tool for BRSs
  BigraphER with new syntax for MDP actions and rewards, better visualisation
  capabilities, and data export options. Furthermore, we introduce a new
  front-end for working with BRSs based on Jupyter notebooks, allowing the user
  to quickly visualise bigraphs and reaction rules, simulate execution, or
  produce a complete transition system, all within the notebook interface.
  Finally, we demonstrate the modelling capabilities of the new system with a
  case study on the movement and behaviour of autonomous agents, exploring ideas
  such as collecting objects, tracking visited locations, hierarchies of spaces,
  and uncertainty in the topology of space.

\keywords{Bigraphs \and Bigraphical reactive systems \and Probabilistic
  semantics \and Discrete-time Markov chain \and Markov decision process \and
  Spatial models.}
\end{abstract}

@@

TODO: DTMC generation, MDP generation: show, define

@@

\begin{definition}[reward structure]
  A \emph{reward structure} for an MDP $(S, \overline{s}, A, P, L)$ is a pair
  $(\rho, \iota)$, where $\rho : S \to \mathbb{R}$ is the \emph{state reward
    function}, and $\iota : S \times A \to \mathbb{R}$ is the \emph{transition
    reward function}.
\end{definition}

@@

% TODO: rework, transition from brs to bigraph
Bigraphical reactive systems (BRS) \cite{DBLP:books/daglib/0022395} are a
universal formalism for modelling interacting systems that evolve in time and
space. In this paper we use a generalised version of bigraphs, known as
\emph{bigraphs with sharing}
\cite{DBLP:phd/ethos/Sevegnani12,dblp:journals/tcs/sevegnanic15}, that uses
directed acyclic graphs instead of forests to represent node placement, allowing
one node to be contained in the intersection of several other nodes. In this
section we introduce the main concepts and definitions in order to expand on
them in Section~\ref{extensions}.

We begin by introducing some notation used in the definitions. Natural numbers
are sometimes interpreted as finite ordinals, i.e., $m = \{ 0, \dots, m - 1 \}$.
We write $S \uplus T$ to indicate the union of sets known or assumed to be
disjoint. For two functions $f$ and $g$ with disjoint domains $S$ and $T$ we
write $f \uplus g$ for the function with domain $S \uplus T$ such that $(f
\uplus g)|_S = f$ and $(f \uplus g)|_T = g$. The identity function on the set
$S$ is written as $\Id_S$. Given a binary relation $\mathit{rel} \subseteq A
\times B$, we denote the domain restriction of $\mathit{rel}$ over $S \subseteq
A$ by $S \lhd \mathit{rel}$. Similarly, we write $\mathit{rel} \rhd S$ for the
range restriction of $\mathit{rel}$ over $S \subseteq B$. We also denote the
transitive closure of $\mathit{rel}$ by $\mathit{rel}^+$. In defining bigraphs
we assume that names, node identifiers, edge identifiers, and controls are drawn
from four infinite pairwise disjoint sets, respectively $\mathcal{X},
\mathcal{V}, \mathcal{E}$, and $\mathcal{K}$.

@@

Informally, a bigraph can be thought of as a graph, where a node can be inside
another node (in this case it can even be in an intersection of several nodes).
However, the pictures tend to become much clearer if we instead visualise the
nesting relations between nodes as just a different type of edges. Consider the
bigraph in Fig.~\ref{example_bigraph}. The white ellipses are the \emph{nodes}.
Each node has a type, called \emph{control}, and denoted here by the labels
$\mathsf{A}, \mathsf{B}$, and $\mathsf{C}$. The green edges are actually
hyperedges (can connect any positive number of nodes), and are called
\emph{links}. In this example they are given names $\mathsf{x}, \mathsf{y}$, and
$\mathsf{e}$. Each node has an ordered list of \emph{ports} (not pictured in
the diagram) that can be thought of as sockets for links. Nodes of the same
control have the same number of ports. The number of ports a node of control
$\mathsf{K}$ can have is called the \emph{arity} of $\mathsf{K}$ and is denoted
by $\ar(\mathsf{K})$. The two white dashed rectangles represent \emph{regions},
and in this representation are the parents to any otherwise parentless nodes.
Grey dashed rectangles are called \emph{sites}, and encode parts of the model
that have been abstracted away. Nodes, sites, and regions are also known as
\emph{places}. The directed black arrows between places show the containment
relations, e.g., the $\mathsf{A}$ node has two children (inner nodes):
$\mathsf{B}$ and the only site in this bigraph, labelled by the number $0$.

Another, more formal way to think about a bigraph is as a pair of graphs: one
for describing the placement of nodes, and one for the links between them. This
is the approach that will be featured in our formal definitions. We also focus
on \emph{concrete} bigraphs, where nodes and links have distinct identifiers.

@@

We define several operations on place graphs that will be used later.

\begin{definition}[composition for place graphs with sharing]
  If $F : k \to m$ and $G : m \to n$ are two concrete place graphs with sharing
  with $V_F \cap V_G = \emptyset$, their composite
  \[ G \circ F = (V, \ctrl, \prnt) : k \to n \]
  has nodes $V \coloneqq V_F \uplus V_G$ and a control map $\ctrl \coloneqq
  \ctrl_F \uplus \ctrl_G$. Its parent relation $\prnt \subseteq (k \uplus V)
  \times (V \uplus n)$ is given by:
  \[ \prnt \coloneqq \prnt_G^\lhd \uplus \prnt_\circ \uplus \prnt_F^\rhd, \]
  where
  \begin{align*}
    \prnt_F^\rhd &= \prnt_F \rhd V_F, \\
    \prnt_G^\lhd &= V_G \lhd \prnt_G, \\
    \prnt_\circ &= (m \lhd \prnt_G) \circ (\prnt_F \rhd m).
  \end{align*}
\end{definition}

\begin{definition}[tensor product for place graphs]
  If 
  \[ G_0 = (V_{G_0}, \ctrl_{G_0}, \prnt_{G_0}) : m_0 \to n_0 \]
  and
  \[ G_1 = (V_{G_1}, \ctrl_{G_1}, \prnt_{G_1}) : m_1 \to n_1 \]
  are two concrete place 
  graphs with sharing with $V_{G_0} \cap V_{G_1} = \emptyset$, their tensor
  product
  \[ G_0 \otimes G_1 = (V, \ctrl, \prnt) : m_0 + m_1 \to n_0 + n_1 \]
  has nodes $V \coloneqq V_{G_0} \uplus V_{G_1}$ and a control map $\ctrl
  \coloneqq \ctrl_{G_0} \uplus \ctrl_{G_1}$. Its parent relation $\prnt
  \subseteq [(m_0 + m_1) \uplus V] \times [V \uplus (n_0 + n_1)]$ is defined as
  \[ \prnt_{G_0} \uplus \prnt_{G_1}^{(m_0, n_0)}, \]
  where
  \begin{align*}
    \prnt_{G_1}^{(m_0, n_0)} = &\{ (v, w) \mid (v, w) \in \prnt_{G_1} \text{ and } v, w \in V_{G_1} \} \\
    \uplus &\{ (m_0 + i, w) \mid (i, w) \in \prnt_{G_1}, w \in V_{G_1} \text{, and } i \in m_1 \} \\
    \uplus &\{ (v, n_0 + i) \mid (v, i) \in \prnt_{G_1}, v \in V_{G_1} \text{, and } i \in n_1 \} \\
    \uplus &\{ (m_0 + i, n_0 + j) \mid (i, j) \in \prnt_{G_1}, i \in m_1 \text{, and } j \in n_1 \}.
  \end{align*}
\end{definition}

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectivelly the \emph{inner} and
  \emph{outer names} of the link graph. $F$ has finite sets $V_F \subset
  \mathcal{V}$ of nodes and $E_F \subset \mathcal{E}$ of edges, a control map
  $\ctrl_F : V_F \to \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

Two places with the same parent, or two points with the same link are called
\emph{siblings}. A link is \emph{idle} if it has no points. Identities over
(inner or outer) name sets are defined by $\id_X = (\emptyset, \emptyset,
\emptyset, \Id_X) : X \to X$.

@@

We use $\epsilon$ as a shorthand for interface $\langle 0, \emptyset \rangle$. A
bigraph with inner face $\epsilon$ is called \emph{ground}.

@@

Given a set of reaction rules, we refer to the configurations that a system may
adopt as \emph{states}. \emph{Rule priorities}
\cite{DBLP:journals/tcs/BaetenBKW89} impose a partial ordering on the reaction
rules, where a rule of lower priority can be applied only if no rule of higher
priority is applicable. Priorities are implemented by assigning each reaction
rule to a \emph{priority class}, where all classes are strictly ordered, and
reaction rules in the same class have the same priority.

\emph{Predicates} can be used to generate labels for states (as in
Definition~\ref{mdp}) or to ease understanding and debugging of the system. They
are expressed in a subset of a logic called BiLog
\cite{DBLP:conf/icalp/ConfortiMS05}, allowing any predicate to be encoded as a
bigraph \cite{DBLP:journals/scp/CalderKSS14,DBLP:phd/ethos/Sevegnani12}.
Therefore, we can check each state for any satisfied predicates by solving a
bigraph matching problem.

\section{Extensions to Bigraphs} \label{extensions}

@@

  Let $g, g'$ be ground bigraphs, and $\{ (r_i, r'_i, p_i) \}_{i=1}^n$ a set of
  ground probabilistic reaction rules, where for each $r_i$, there exists a
  bigraph $D_i$ such that $g = D_i \circ r_i$. Let $S = \{ (r_i, r'_i, p_i) \mid
  g' = D_i \circ r'_i \}$ (for the same $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation is defined as
  \[ g \xrightarrowrhd{p}_{\mathcal{R}} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', p') \in S} p'. \]
\end{definition}

@@

  Let $g, g'$ be ground bigraphs, $a \in A$ an action, and $\{ (r_i, r'_i, a,
  p_i) \}_{i=1}^n$ a set of ground nondeterministic reaction rules with action
  $a$, where for each $r_i$, there exists a bigraph $D_i$ such that $g =
  D_i \circ r_i$. Let $S = \{ (r_i, r'_i, a, p_i) \mid g' = D_i \circ r'_i \}$
  (for the same $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation for action $a$ is defined as
  \[ g \xrightarrowrhd[r(a)]{p}_{a} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', a, p') \in S} p'. \]
\end{definition}

Finally, we can associate a reward with each predicate, allowing us to assign
rewards to states in a flexible and semantically meaningful way: the reward of a
state is simply the sum of the rewards of all matching predicates (and $0$ in
case there are none).

@@

TODO: introduce BigraphER \cite{DBLP:conf/cav/SevegnaniC16}

@@

The MDP can be represented as an NBRS with BigraphER code in
Listing~\ref{example_listing}. Furthermore, the BigraphER's visualisation of the
(automatically generated) transition system is in Fig.~\ref{example_ts}.

@@

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. Text inside state ellipses lists all predicates that are satisfied by
that state. Edges from actions to states are labelled with both the probability
and the name of the reaction rule. Transitions from either a fully generated
transition system or a simulation, labels from predicates, and state/transition
rewards can be exported to the various plain text formats of a probabilistic
model checker PRISM \cite{DBLP:conf/cav/KwiatkowskaNP11} for further
quantitative analysis.

TODO: introduce jupyter (with more/new screenshots)

\begin{figure}
  \centering
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

@@

TODO: Akin to \cite{dblp:conf/nfm/giaquintahimn18} and maybe
\cite{koenig1998xavier}.

@@

TODO: explain the syntax of the language

Our first scenario is concerned with collecting objects in a grid-like
two-dimensional world. The agent starts in the southwest corner of the map
(which we call \emph{home}), explores the environment until it collects a
specified number of objects, and returns home (while possibly collecting more
objects along the way). The model tracks which cells of the grid have been
explored, and only the unexplored cells have a constant probability $p$ of
rewarding the agent with an object. The desired number of objects can be easily
changed, and the map can be customised to make some transitions unavailable
(i.e. add walls), as long as the agent can still return home.

@@

We begin with a list of controls in Listing~\ref{agent1_controls}. On the right
hand side of each equal sign is the number of ports that each node of that
control has, while the \texttt{\textbf{atomic}} keyword specifies the nodes that
cannot contain other nodes. \texttt{Cell}s represents discrete locations. They
are linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all
except one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}. In order to represent multiple
objects, it is convenient to put the second \texttt{Object} inside the first,
and so on. \texttt{Directions} contains a subset of the four nodes,
corresponding to the four possible directions of travel: \texttt{North},
\texttt{East}, \texttt{West}, \texttt{South}. We connect neighbouring
\texttt{Cell}s by linking the \texttt{West} node of one \texttt{Cell} to the
\texttt{East} node of another (or \texttt{North} to \texttt{South}).

@@

Next, we define two predicates: \texttt{home} and \texttt{goal} (see
Listing~\ref{agent1_code}). The former states that the agent is home if it is in
the southwest corner of the grid, while the latter specifies a goal of
collecting one object. A larger number of objects can be set by nesting
\texttt{Object}s inside each other, e.g., \texttt{Agent.Object.Object}.
Additionally, Listing~\ref{agent1_code} defines the initial $2 \times 2$ grid,
as described previously.

@@

Then, we have an action for each of the four directions, each with three
reaction rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the first time and
finding an object. We will use the rules for going north in
Listing~\ref{agent1_north} as an example. When going to an already visited cell,
the \texttt{Agent} simply traverses a link between the \texttt{North} node of
its previous \texttt{Cell} and the corresponding \texttt{South} node. As this is
the only reaction rule for this situation, its probability is $1$. When going to
an unvisited node, we collect an \texttt{Object} with probability $p$, or make
the transition without getting an \texttt{Object} with probability $1-p$.
Regardless, the \texttt{Cell} changes its link from \texttt{Unvisited} to
\texttt{Visited}.

@@

Lastly, we have three actions for when the agent acquires the specified number
of objects: two for going home west or south (since these are the two directions
necessary and sufficient to reach home), and one for staying home. The two
actions for going home have two reaction rules each, depending on whether the
destination cell is visited or not. They are just like the rules in
Listing~\ref{agent1_north}, except every mention of \texttt{Agent} is replaced
with \texttt{goal}. The rule to stay home simply says that if the agent has
achieved the goal, and is home (characterised by a \texttt{Cell} with only two
directions: \texttt{North} and \texttt{East}), then nothing should change (see
Listing~\ref{agent1_home}). To make the agent start heading home as soon as
enough objects are collected, all five reaction rules related to going and
staying home are in a higher priority class.

@@

TODO: Should mention \cite{DBLP:conf/giscience/WaltonW12}.

@@

In order to impement nesting, a \texttt{Cell} contains a \texttt{Node}, which
can have any number of \texttt{Cell}s of its own. Since after entering a room,
you end up in a specific place inside the room (next to the door), we also need
to link up the \texttt{Node} with one of the \texttt{Cell}s it contains (see
Fig.~\ref{agent2_nesting}). Both \texttt{Cell}s and \texttt{Node}s support
having a single link for exactly this purpose. Thus, most of the \texttt{Cell}
ports are not linked to anything, which is implemented via closures (e.g.
\texttt{/x Cell\{x\}}). The initial map in Listing~\ref{agent2_initial} has a
corridor of three cells, with the agent on the west-most cell, and a door to a
room in the middle cell. One moves in and out of a room via a pair of actions
with reaction rules pictured in Fig.~\ref{go_in_out}.

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1})\footnote{Multiple different rooms can be implemented by
having each previously empty \texttt{Node} contain a node of different control.
This way a different probability distribution (with different possible outcomes)
can be assigned to each room.}. In order for this to happen before any other
action can take place, the two reaction rules in Listing~\ref{agent2_room} are
put on a higher priority class. With probability $p$, the generated room has
each \texttt{Cell} linked to its two neighbours; and with probability $1-p$, the
top two \texttt{Cell}s are not linked, possibly forcing the agent to choose a
different path. Finally, we have four actions, each with a single reaction rule
with probability $1$, corresponding to the four possible directions of travel.
The reaction rules are very similar to those of our previous example, see
Listing~\ref{agent2_north}.
